---
title: "Stoke Data"
author: "Miles Tweed"
date: "5/11/2021"
output: pdf_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(olsrr)
library(car)
```

```{r, message=FALSE}
Stroke <- read_csv('healthcare-dataset-stroke-data.csv')
Stroke$bmi <- Stroke$bmi %>% as.numeric()
Stroke <- Stroke %>% mutate(bmi2 = ifelse(is.na(bmi), median(bmi, na.rm = TRUE), bmi)) %>% select(-bmi, -id) %>% filter(gender!="Other")
Stroke$stroke <- factor(Stroke$stroke, levels=c(0,1), labels = c("No","Yes"))
```

# Expectation 1

```{r}
ggplot(Stroke, aes(x = gender, fill =  gender)) + 
  geom_bar(position = 'dodge') +
  labs(title = "Gender Distribution")

ggplot(Stroke, aes(x = work_type, fill =  gender)) + 
  geom_bar(position = 'dodge')

ggplot(Stroke, aes(x = Residence_type, fill =  gender)) + 
  geom_bar(position = 'dodge')

ggplot(Stroke, aes(x = smoking_status, fill =  gender)) + 
  geom_bar(position = 'dodge')

ggplot(Stroke, aes(x = ever_married, fill =  gender)) + 
  geom_bar(position = 'dodge')

ggplot(Stroke, aes(x = stroke, fill = stroke)) + 
  geom_bar(position = 'dodge')
```


# Expectation 2
```{r}
# Making ever_married and smoking_status factors
Stroke$ever_married <- factor(Stroke$ever_married)
Stroke$smoking_status <- factor(Stroke$smoking_status)

# Contingency Table
con.table <- table(Stroke$ever_married, Stroke$smoking_status)
con.table

# Chi-Squared Test
chisq.test(con.table)

# Probability that marriage and smoking status are independent
pchisq(600.33, df=3, lower.tail=F)

x11()
curve(dchisq(x, df = 3), from = 0, to = 20, main = "Chi-Squared Distribution with df=3")
savePlot("chi_sq.png")
curve(dchisq(x, df = 3), from = 0, to = 601, main = "Location of Test Statistic")
abline(v = 600.33, col='red')
text(x=530, y = 0.03, labels = "Chi-Squared Statistic", col = 'red', cex = 0.7)
savePlot("chi_sq_stat.png")
```



# Expectation 3
```{r}
lm.obj <- lm(bmi2~., Stroke)

summary(lm.obj)

step(lm.obj, trace = 0)

lm.reduced <- lm(bmi2 ~ age + hypertension + heart_disease + ever_married + 
    work_type + avg_glucose_level + smoking_status + stroke, data = Stroke)

summary(lm.reduced)
```


# Expectation 4
```{r}
plot(lm.reduced,which=1)
plot(lm.reduced,which=2)
```

# Expectation 5

```{r}
plot(lm.reduced, which=4)

# remove outliers based on Cook's distance
Stroke.Out <- Stroke[-c(2129, 4209, 4611),]

# Refit Model
lm.outliers <- lm(bmi2~age + hypertension + heart_disease + ever_married + 
    work_type + avg_glucose_level + smoking_status + stroke, data = Stroke.Out)
summary(lm.outliers)
```


# Expectation 6
```{r}
library(car)
glm.obj <- glm(stroke~., Stroke, family='binomial')

glm.null <- glm(stroke~1, Stroke, family = 'binomial')

vif(glm.obj)

# Test of overall model significance
# Likelihood Ratio Test
anova(glm.null, glm.obj, test = "LRT")
Anova(glm.obj)
```

```{r}
# Backwards Stepwise Selection
step(glm.obj, trace = 0)
```

Let $p_i = p(stroke_i = 1\ |\ age_i,\ hypertension_i,\ heart\_disease_i,\  avg\_glucose\_level_i)$

$$\left\{\begin{array}{rl}stroke_i&\sim_{indep.}Bin(1,p_i),\\\log\left(\frac{p_i}{1-p_i}\right)&=\beta_0 +\beta_1age_i+\beta_2D_{hypertension ,i}+\\&\beta_3D_{heart_disease,i}+ \beta_4avg\_glucose\_level_i\end{array}\right.$$

```{r}
# Fit the reduced model
glm.reduced <- glm(stroke ~ age + hypertension + heart_disease + avg_glucose_level, data = Stroke, family = 'binomial')
glm.reduced$coefficients
```

\begin{align*}
\log\left(\frac{\hat{p}_i}{1-\hat{p}_i}\right)&=-7.489 +0.0689age_i+0.381D_{hypertension ,i}+\\&0.330D_{heart_disease,i}+ 0.004avg\_glucose\_level_i
\end{align*}


```{r}

#checking for collinearity
vif(glm.reduced)

anova(glm.null, glm.reduced, test = "LRT")

# Reduced model accuracy
glm.pred <- predict(glm.reduced, type='response')
stroke.pred <- ifelse(glm.pred > 0.50, "Yes","No")
stroke.labs <- Stroke$stroke

# Confusion matrix
conf.mat <- table(Pred=stroke.pred, 
                  True=stroke.labs)
conf.mat

# Misclassification rate
mean(stroke.pred != stroke.labs)



glm.pred <- predict(glm.obj, type='response')
stroke.pred <- ifelse(glm.pred > 0.50, "Yes","No")
stroke.labs <- Stroke$stroke

# Confusion matrix
conf.mat <- table(Pred=stroke.pred, 
                  True=stroke.labs)
conf.mat

# Misclassification rate
mean(stroke.pred != stroke.labs)

Confint(glm.reduced)
```


```{r}

ggplot(Stroke, aes(x = stroke, fill = stroke)) + 
  geom_bar(position = 'dodge') + 
  labs(title="Class Distribution") +
  ggsave("classDist.png", width = 100, height = 60, units = 'mm')

library(ROSE)

Stroke.bal <- ovun.sample(stroke~., data = Stroke)$data

ggplot(Stroke.bal, aes(x = stroke, fill = stroke)) + 
  geom_bar(position = 'dodge') + 
  labs(title="Class Distribution") +
  ggsave("classDistBal.png", width = 100, height = 60, units = 'mm')


glm.obj.bal <- glm(stroke~., Stroke.bal, family='binomial')

glm.null.bal <- glm(stroke~1, Stroke.bal, family = 'binomial')

vif(glm.obj.bal)
# Test of overall model significance
# Likelihood Ratio Test
anova(glm.null.bal, glm.obj.bal, test = "LRT")

# Forward Stepwise Selection
step(glm.obj.bal)

# Fit the reduced model
glm.reduced.bal <-glm(formula = stroke ~ age + hypertension + heart_disease +
                      work_type + Residence_type + avg_glucose_level +
                      smoking_status, family = "binomial", data = Stroke.bal)

glm.reduced.bal$coefficients
```
\begin{align*}
\text{Let }p_i&= p(stroke_i = 1\ |\ age_i,\ hypertension_i,\ heart\_disease_i,\\
&\ work\_type_i,\ Residence\_type_i,\ avg\_glucose\_level_i,\ smoking\_status_i)
\end{align*}

$$\left\{\begin{array}{rl}stroke_i&\sim_{indep.}Bin(1,p_i),\\
\log\left(\frac{p_i}{1-p_i}\right)&=\beta_0 +\beta_1age_i+\beta_2D_{hypertension ,i}+\beta_3D_{heart\_disease,i}\\
&+\beta_4D_{Govt\_job,i}+\beta_5D_{Never\_worked,i}+\beta_6D_{Private,i}+\beta_7D_{Self-employed,i}\\
&+\beta_8D_{Urban,i}+\beta_9avg\_glucose\_level_i+\beta_{10}D_{Never\_Smoked,i}\\
&+\beta_{11}D_{Smokes,i}+\beta_{12}D_{SmokesUNK,i}
\end{array}\right.$$

\begin{align*}
\log\left(\frac{\hat{p}_i}{1-\hat{p}_i}\right)&=-4.367 +0.082age+0.29D_{hypertension}+0.22D_{heart\_disease,i}\\
&-0.92D_{Govt\_job,i}-11.02D_{Never\_worked,i}-1.02D_{Private,i}-1.13D_{Self-employed,i}\\
&+0.199D_{Urban,i}+0.005avg\_glucose\_level_i-0.29D_{Never\_Smoked,i}\\
&+0.099D_{Smokes,i}-0.77D_{SmokesUNK,i}
\end{align*}

```{r}
vif(glm.reduced.bal)

anova(glm.null.bal, glm.reduced.bal, test = "LRT")

Anova(glm.reduced.bal)

Confint(glm.reduced.bal)


glm.pred.bal <- predict(glm.reduced.bal, type='response')
stroke.pred.bal <- ifelse(glm.pred.bal > 0.50, "Yes","No")
stroke.labs.bal <- Stroke.bal$stroke

# Confusion matrix
conf.mat.bal <- table(Pred=stroke.pred.bal, 
                  True=stroke.labs.bal)
conf.mat.bal

# Misclassification rate
mean(stroke.pred.bal != stroke.labs.bal)

```